---
title: "Homework 2"
author: "Mari Sanders"
output: github_document
---
```{r message=FALSE, }
library(tidyverse)
library(readxl)
options(scipen=999)
```

# Problem 1 

```{r}
transit_df <- read_csv("data_hw2/nyc_subway_data.csv") %>% 
  janitor::clean_names() %>% 
  select(
    -c("division", "entrance_type", "exit_only", "staffing", "staff_hours",
       "ada_notes", "free_crossover", "north_south_street", "east_west_street", 
       "corner", "entrance_latitude", "entrance_longitude", "station_location", 
       "entrance_location")
  ) %>% 
  mutate(
    entry = case_match(
      entry, 
      "YES" ~ TRUE, 
      "NO" ~ FALSE
    )
  ) 
```

The data contains the variables `r names(transit_df)`. To clean the data, I started by using the `clean_names()` function from the `janitor` package. Then I selected only the rows that were needed for the analysis by using `select()`. Next, I changed the entry variable to a logical variable instead of a character variable. This is fairly tidy, except the routes that each station serve it a bit hard to understand. This cleaned data contains `r nrow(transit_df)` rows and `r ncol(transit_df)` columns. 

```{r}
distinct_station <- 
  transit_df %>%
  distinct(line, station_name) %>%
  nrow()
```

There are `r distinct_station` distinct stations in this dataset. 

```{r}
ada_compliant <- transit_df %>% 
  filter(ada == "TRUE") %>% 
  distinct(line, station_name) %>% 
  nrow()
```

There are `r ada_compliant` ada compliant stations. 

```{r}
clean_transit <- transit_df %>% 
  mutate(
    across(starts_with("route"), 
           as.character)
  ) %>% 
  pivot_longer(
    cols = route1:route11,
    names_to = "route_number",
    values_to = "route_name"
  ) 
```

```{r}
a_stations <- clean_transit %>% 
  filter(route_name == "A") %>%  
  distinct(line, station_name) %>%  
  nrow()
```

There are `r a_stations` distinct stations that serve the A train

```{r}
a_ada_stations <- 
  clean_transit %>% 
  filter(route_name == "A" & ada == "TRUE") %>%  
  distinct(line, station_name) %>%  
  nrow()
```

There are `r a_ada_stations` stations that serve the a train and are ada accessible. 

# Problem 2 

```{r}
mr_trash_wheel_df <- read_excel("data_hw2/202309_Trash_Wheel_Collection_Data.xlsx", 
                                range = "A2:N586", sheet = "Mr. Trash Wheel") %>% 
  janitor::clean_names() %>% 
  select(
    -c("homes_powered")
  ) %>% 
  mutate(
    sports_balls, sports_balls = as.integer(sports_balls)) %>% 
  mutate(trash_wheel = "mr.") %>% 
  mutate(year, year = as.numeric(year))


prof_trash_wheel_df <- 
  read_excel("data_hw2/202309_Trash_Wheel_Collection_Data.xlsx", 
             range = "A2:M108", sheet = "Professor Trash Wheel") %>% 
  janitor::clean_names() %>% 
  select(
    -c( "homes_powered")
  ) %>% 
  mutate(trash_wheel = "professor") %>% 
  mutate(year, year = as.numeric(year))

gwynnda_df <- 
  read_excel("data_hw2/202309_Trash_Wheel_Collection_Data.xlsx", 
             range = "A2:L157", sheet = "Gwynnda Trash Wheel") %>% 
  janitor::clean_names() %>% 
  select(
    -c("homes_powered")) %>% 
  mutate(
    trash_wheel = "gwynnda"
  ) %>% 
  mutate(year,year = as.numeric(year))

trash_wheel_df <- 
  bind_rows(mr_trash_wheel_df, prof_trash_wheel_df,gwynnda_df) %>% 
  relocate(trash_wheel, .before = month)
```

This data contains variables `r names(trash_wheel_df)`. To clean each of the data sets, I used `clean_names()` from the package `janitor`, then I selected only the variables that I needed, which were all the columns that contained information about the dumpster-specific information. Finally, to keep track of which trash wheel the data is from, I created a variable called `trash_wheel` that contained values `mr.`, `gwynnda`, or `professor`. There are `r nrow(trash_wheel_df)` observations and `r ncol(trash_wheel_df)` variables. 

```{r}
total_weight_prof <- 
  trash_wheel_df %>% 
  filter(trash_wheel == "professor") %>% 
  pull(weight_tons) %>% 
  sum()
```

The total weight collected by Professor Trash Wheel is `r total_weight_prof` tons. 

```{r}
cig_butts_gwynnda <- 
  trash_wheel_df %>% 
  filter(trash_wheel == "gwynnda") %>% 
  filter(year == 2022) %>% 
  filter(month == "June") %>% 
  pull(cigarette_butts) %>% 
  sum()
```

Gwynnda collected `r cig_butts_gwynnda` cigarette butts in June 2022. 

## Problem 3 

```{r}
baker_df <- 
  read_csv("data_hw2/bakers.csv") %>% 
  janitor::clean_names()

bakes_df <- 
  read_csv("data_hw2/bakes.csv") %>%
  janitor::clean_names()

results_df <- 
  read_csv("data_hw2/viewers.csv") %>% 
  janitor::clean_names() 
```


