---
title: "Homework 2"
author: "Mari Sanders"
output: github_document
---
```{r message=FALSE, }
library(tidyverse)
library(readxl)
options(scipen = 999)
```

# Problem 1 

```{r}
transit_df <- read_csv("data_hw2/nyc_subway_data.csv") %>% 
  janitor::clean_names() %>% 
  select(
    -c("division", "entrance_type", "exit_only", "staffing", "staff_hours",
       "ada_notes", "free_crossover", "north_south_street", "east_west_street", 
       "corner", "entrance_latitude", "entrance_longitude", "station_location", 
       "entrance_location")
  ) %>% 
  mutate(
    entry = case_match(
      entry, 
      "YES" ~ TRUE, 
      "NO" ~ FALSE
    )
  ) 
```

The data contains the variables `r names(transit_df)`. To clean the data, I started by using the `clean_names()` function from the `janitor` package. Then I selected only the rows that were needed for the analysis by using `select()`. Next, I changed the entry variable to a logical variable instead of a character variable. This is fairly tidy, except the routes that each station serve it a bit hard to understand. This cleaned data contains `r nrow(transit_df)` rows and `r ncol(transit_df)` columns. 

```{r}
distinct_station <- 
  transit_df %>%
  distinct(line, station_name) %>%
  nrow()
```

There are `r distinct_station` distinct stations in this dataset. 

```{r}
ada_compliant <- transit_df %>% 
  filter(ada == "TRUE") %>% 
  distinct(line, station_name) %>% 
  nrow()
```

There are `r ada_compliant` ada compliant stations. 

```{r}
clean_transit <- transit_df %>% 
  mutate(
    across(starts_with("route"), 
           as.character)
  ) %>% 
  pivot_longer(
    cols = route1:route11,
    names_to = "route_number",
    values_to = "route_name"
  ) 
```

```{r}
a_stations <- clean_transit %>% 
  filter(route_name == "A") %>%  
  distinct(line, station_name) %>%  
  nrow()
```

There are `r a_stations` distinct stations that serve the A train

```{r}
a_ada_stations <- 
  clean_transit %>% 
  filter(route_name == "A" & ada == "TRUE") %>%  
  distinct(line, station_name) %>%  
  nrow()
```

There are `r a_ada_stations` stations that serve the a train and are ada accessible. 

# Problem 2 

```{r}
mr_trash_wheel_df <-read_excel("data_hw2/202409_Trash_Wheel_Collection_Data.xlsx",range = "A2:N653", sheet = "Mr. Trash Wheel") %>% 
  janitor::clean_names() %>% 
  select(
    -c("homes_powered")
  ) %>% 
  mutate(
    sports_balls, sports_balls = as.integer(sports_balls)) %>% 
  mutate(trash_wheel = "mr.") %>% 
  mutate(year, year = as.numeric(year))


prof_trash_wheel_df <- 
  read_excel("data_hw2/202409_Trash_Wheel_Collection_Data.xlsx", 
             range = "A2:M121", sheet = "Professor Trash Wheel") %>% 
  janitor::clean_names() %>% 
  select(
    -c( "homes_powered")
  ) %>% 
  mutate(trash_wheel = "professor") %>% 
  mutate(year, year = as.numeric(year))

gwynnda_df <- 
  read_excel("data_hw2/202409_Trash_Wheel_Collection_Data.xlsx", 
             range = "A2:L265", sheet = "Gwynnda Trash Wheel") %>% 
  janitor::clean_names() %>% 
  select(
    -c("homes_powered")) %>% 
  mutate(
    trash_wheel = "gwynnda"
  ) %>% 
  mutate(year,year = as.numeric(year))

trash_wheel_df <- 
  bind_rows(mr_trash_wheel_df, prof_trash_wheel_df,gwynnda_df) %>% 
  relocate(trash_wheel, .after = month)
```

This data contains variables `r names(trash_wheel_df)`. To clean each of the data sets, I used `clean_names()` from the package `janitor`, then I selected only the variables that I needed, which were all the columns that contained information about the dumpster-specific information. Finally, to keep track of which trash wheel the data is from, I created a variable called `trash_wheel` that contained values `mr.`, `gwynnda`, or `professor`. There are `r nrow(trash_wheel_df)` observations and `r ncol(trash_wheel_df)` variables. 

```{r}
total_weight_prof <- 
  trash_wheel_df %>% 
  filter(trash_wheel == "professor") %>% 
  pull(weight_tons) %>% 
  sum()
```

The total weight collected by Professor Trash Wheel is `r total_weight_prof` tons. 

```{r}
cig_butts_gwynnda <- 
  trash_wheel_df %>% 
  filter(trash_wheel == "gwynnda") %>% 
  filter(year == 2022) %>% 
  filter(month == "June") %>% 
  pull(cigarette_butts) %>% 
  sum()
```

Gwynnda collected `r cig_butts_gwynnda` cigarette butts in June 2022. 

## Problem 3 

```{r}

baker_df <- 
  read_csv("data_hw2/bakers.csv") %>% 
  janitor::clean_names() %>% 
  separate(baker_name, into = c("first_name", "last_name"), sep = " ")

bakes_df <- 
  read_csv("data_hw2/bakes.csv") %>%
  janitor::clean_names() 

results_df <- 
  read_csv("data_hw2/results.csv", skip = 2) %>% 
  janitor::clean_names()

final_data <- baker_df %>% 
  left_join(bakes_df, by = c(first_name = "baker", series = "series")) %>% 
  left_join(results_df, by = c(series = "series", episode = "episode", first_name = "baker")) %>% 
  relocate(series, episode, .before = "first_name")

write_csv(final_data, "ggb_data.csv")
```


The first thing that I did with all of the datasets is I used the `clean_names()` function from the `janitor` package to make all the names of the columns easier to work with. For the `baker_df()`, I separated the `baker_name` into `first_name` and `last_name`, so that it would be easier to join the data later. To combine each of the data, I first used `left_join` by `first_name` and `series` because I wanted to have all of the data from `baker_df` to be included. Then I used `left_join` to join the `results_df` with the rest of the combined data by `series`, `episode` and `first_name`. Finally, to make the data more readable, I relocated `series` and `episode` to be before `first_name`. This dataset includes the variables `r names(final_data)`. It has `r ncol(final_data)` columns and `r nrow(final_data)` rows. This dataset is still a bit messy because in order to show each of the bakers `show_stopper` and `signature_bake`, the data about their  `series`, `episode`, `first_name`, `last_name`, `baker_age`, `baker_occupation`, and `hometown` are duplicated for the amount of times they baked a signature_bake and a show_stopper. I think it would be best to keep the data seperate and then use the data that you need from both of them, without combining them. 

```{r}
final_data %>% 
  filter(series > 5) %>% 
    filter(result == "STAR BAKER" | result == "WINNER") %>% 
  knitr::kable(., col.names = c("Season", "Episode", "First Name", "Last Name", "Age", "Occupation", "Hometown", "Signature Bake", "Show Stopper", "Technical Score", "Result")) 
```



```{r}
viewership_df <- 
  read_csv("data_hw2/viewers.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    cols = series_1:series_10,
    names_to = "season", 
    names_prefix = "series_"
  ) %>% 
  mutate(season, season = as.numeric(season)) %>% 
  arrange(season) %>% 
  relocate(season, .before = episode) 

head(viewership_df, 10)

avg_view_1 <- 
  viewership_df %>% 
  filter(season == 1) %>% 
  summarise(avg = mean(value, na.rm = TRUE))  %>% 
  pull()

avg_view_5 <- 
  viewership_df %>% 
  filter(season == 5) %>% 
  summarise(avg = mean(value, na.rm = TRUE))  %>% 
  pull()
``` 

The average viewership in Season 1 was `r avg_view_1`. The average viewership in Season 5 was `r avg_view_5`. 